Synthesizing Symmetric Lenses

### 2019-03-01 papers/SymmetricLenses/main.tex
add updated info

In summary, our contributions are as follows:
\begin{itemize}
%
\item We define {\em simple symmetric lenses}, a subclass of symmetric
lenses with no hidden state,
%
we propose a syntax for them---a collection
of combinators for building simple symmetric lenses---and we
describe how to calculate their likelihoods
(\S\ref{sec:overview}--\S\ref{sec:ssl}).  
\item We give an algorithm for synthesizing lenses built from these combinators,
  using a novel application of stochastic regular expressions to guide the
  search process (\S\ref{sec:synthesis}) and prove that our metrics have deep
  semantic meaning. \bcp{``Semantic meaning'' is redundant.  And ``deep
    semantic meaning'' just sounds silly.  Can we say it more precisely?}
%
\item We extend the Boomerang implementation with simple symmetric lenses
and lens synthesis specifications, 
  including types, examples, and relevance annotations. We evaluate our
  implementation, the effects of
  optimizations, and our inclusion of relevance annotations on a set of
  48 lens synthesis benchmarks drawn from data cleaning, view maintenance, and
  file synchronization tasks. The system can synthesize simple symmetric lenses
  for all of the benchmarks in under 30 seconds (\S\ref{sec:evaluation}).
%
\item We prove that simple symmetric lenses fall between asymmetric
lenses~\cite{Focal2005-long} and full symmetric
lenses~\cite{symmetric-lenses} in expressiveness.  Indeed, the
class of simple lenses can be characterized semantically as the subset of
full symmetric lenses for which a simple ``forgetfulness'' property holds
(\S\ref{sec:related}).
\end{itemize}
Along the
way, we show that {\em star-semiring equivalences}\bcp{citation} on regular
expressions 
remain valid when extended to stochastic regular expressions, preserving
probability distributions as well as languages (\S\ref{sec:sre}).
We close with related work (\S\ref{sec:related}) and
concluding thoughts (\S\ref{sec:conc}).
Elided proofs can be found in the appendices
\ifappendices\else included with the supplemental material\fi.

@@

\subsection{Simple Symmetric Lenses}
Semantically, a simple symmetric lens (we will just say ``lens'' when clear
from context) between 
sets $X$ and $Y$ comprises four functions subject to four round-tripping
laws.
\vspace*{.7ex}
\begin{trivlist}
\item 
%\begin{enumerate}
%\item $\CreateR \OfType \Regex \rightarrow \RegexAlt$
%\item $\CreateL \OfType \RegexAlt \rightarrow \Regex$
%\item $\PutR \OfType \Regex \rightarrow \RegexAlt \rightarrow \RegexAlt$
%\item $\PutL \OfType \RegexAlt \rightarrow \Regex \rightarrow \Regex$
%\end{enumerate}
\begin{center}
\begin{minipage}[c]{.3\textwidth}
\begin{gather}
 \mbox{$\CreateR \OfType X \rightarrow Y$} \nonumber \\
 \mbox{$\CreateL \OfType Y \rightarrow X$} \nonumber \\
 \mbox{$\PutR \OfType X \rightarrow Y \rightarrow Y$} \nonumber \\
 \mbox{$\PutL \OfType Y \rightarrow X \rightarrow X$} \nonumber
\end{gather} 
\end{minipage} 
\hspace{.4in}
\begin{minipage}[c]{.5\textwidth}
\begin{align}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{x})}{x} = x\\
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{y})}{y} = y\\
  \tag{\PutRL}
  \PutLOf{(\PutROf{x}{y})}{x} = x\\
  \tag{\PutLR}
  \PutROf{(\PutLOf{y}{x})}{y} = y
\end{align}
\end{minipage} 
\end{center}
\end{trivlist}
The two \KW{create} functions are used to \proposedelete{propagate shared
  information and} fill 
in default values when introducing new data (\EG, on \KW{create} the ``unk''
salary entry is added alongside the name to the management file when HR inserts
a new employee). The two \KW{put} functions propagate edits from one format to
the other by combining a new value from one with an old value from the other.
The record projection notation $\Lens.\PutR$ extracts the \PutR function
from the 
lens $\Lens$. These four functions can be used to keep the
common information between two file types in sync. For example, if a new file
of the left-hand format is created, the \CreateR function will build a new
file in the 
right-hand format. If the right-hand file is then edited, the \PutL function can
update the left-hand file with the changed information.

@@

% concat(ins(header),employees_lens) : emp_salaries $\Leftrightarrow$ emp_insurance
%
%\ksf{I don't think this example is necessary.}
%Data can also appear in multiple forms. For example,
%a list of employees with insurance information can either be empty or
%be a sequence of employee/insurance records: \ksf{This example is
%vacuous as you could eliminate the or in favor of the *.  Do somethign
%else to introduce or?}
%
%\begin{lstlisting}
%or(id(""),employees_lens) : emp_salaries $\Leftrightarrow$ "" | ("\n" emp_ins) . ("\n" emp_ins)*
%\end{lstlisting}
The \IterateLens lens is useful for synchronizing a series of items or rows in a
table. For example given a
lens \lstinline{employee_lens} that synchronizes data for a single employee, the
lens\bcp{In all these examples, I'd like to have one more space on each side
of the ``:''}
%
\begin{lstlisting}
iterate(id("\n") . employee_lens) : ("\n" . emp_salary)$^*$ $\Leftrightarrow$ ("\n" . emp_ins)$^*$
\end{lstlisting}
%
transforms a list of employees in employees in the Management format to a list
of employees in the HR format and vice versa. These combinators
are combined in figure~\ref{fig:example_lens} to construct a complete lens
between the employee formats.

@@

\bcp{I wonder whether the rest of this subsection can be cut... It is both
  hard to understand and too brief to really say much, and I wonder whether
  we need it for anything between here and where we give it a real
  explanation.}

Most of the typing rules for lenses are syntax directed, where only a handful of
rules result in a given type\bcp{syntax directed usually means that the
  conclusion of at most
  {\em one} rule will match a given goal; if that's not the case here, then
  we need to speak more carefully}.  For example, the type of a
\ConcatLens lens is the concatenation of the regular expressions of the sub-lenses.
\[
  \inferrule*
  {
    \Lens_1 \OfType \BRegex_1 \Leftrightarrow \BRegexAlt_1\\
    \Lens_2 \OfType \BRegex_2 \Leftrightarrow \BRegexAlt_2
  }
  {
    \ConcatLensOf{\Lens_1}{\Lens_2} \OfType \BRegex_1 \Concat \BRegex_2
    \Leftrightarrow
    \BRegexAlt_1 \Concat \BRegexAlt_2
  }
\]
The typing  rules for maintain structural similarity
between regular expression types and lens terms, with one exception: The
type equivalence rule permits a well-typed lens to be well-typed among any
equivalent regular expression pair.
%
\[
  \centering
  \inferrule*
  {
    \Lens \OfType \BRegex \Leftrightarrow \BRegexAlt\\
    \BRegex \SSREquiv \BRegex'\\
    \BRegexAlt \SSREquiv \BRegexAlt'
  }
  {
    \Lens \OfType \BRegex' \Leftrightarrow \BRegexAlt'
  }
\]
%
This rule is needed for lenses like the final one in
Figure~\ref{fig:example_lens} to be well-typed.

@@

One challenge is that the simple symmetric lens combinators permit many
well-typed lenses
between a given pair of regular expressions. For example, 
Figure~\ref{fig:example_lens} gives one possible lens between regular
expressions  \lstinline{emp_salaries} and \lstinline{emp_insurance},
but 
%
\begin{lstlisting}
disconnect(emp_salaries, emp_insurance, ..., ...)
\end{lstlisting}
%
is another well-typed lens that satisfies the example in
Figure~\ref{fig:minimized-representations} (where \lstinline{...} represents
components of the example itself\bcp{huh?}). In general, {\em many} examples
may be 
required to rule out all possible occurrences of nested\bcp{??} \Disconnect lenses,
particularly in complex formats.
%\ksf{I don't understand the following.  We shouldn't explain the issue in
%terms of \PutR and \PutL as we are focused on the combinators rather
%than the raw functions.}
%The default strings given to \Disconnect merely need to be the strings
%in the input-output examples, and the lens will still satisfy the
%specification.  While the disconnect lens is prohibited by two
%examples, or by a carefully chosen input-output example on \PutR
%or \PutL, this problem continues presenting itself for the sublenses,
%and is highlighted even more on larger formats.
\proposedelete{While we could ask programmers to provide detailed and precise specifications
for their synthesis tasks, we would prefer an algorithm that can intelligently
choose a lens despite the underspecification.} Instead of merely finding
\emph{any} satisfying lens, we wish to synthesize a satisfying lens that is
likely to please the user.

How can we identify such a ``likely'' lens? We propose the following heuristic:
A satisfying lens is ``more likely'' if it uses more data from one format to
construct the other.  For example, the
identity lens (which uses all the data) is more likely than the disconnect
lens (which uses none). Formally, we define the {\em likelihood} of a
satisfying lens as the 
expected number of bits required to recover data in one format from 
data in the other; higher likelihoods correspond to fewer bits. Two
strings $s$ and $t$ are \emph{synchronized according to lens
  $l$}\bcp{Earlier we used $\Lens$, now we are switching to $l$...} if $l.\PutR\
s\ t = t$ and $l.\PutL\ t\ s = s$. We can \emph{recover} $s$ from 
$t$ using bits $b$ and lens $l$ if we can reconstruct $s$ from $t$, $b$, and
$l$. For example, given the \IdentityLens{} lens, we can recover $s$ from $t$
using no bits because, in this case, $s$ is just $t$. In contrast, given the
\Disconnect{} lens, we need enough bits to fully encode $s$ in order to recover
it from $t$ because all the information in $t$ gets thrown away. Thus, if both
\IdentityLens{} and \Disconnect{} are satisfying lenses for a particular pair of
formats, \IdentityLens{} will be more likely.

@@

We use entropy to gauge the relative likelihood of lenses in our synthesis
algorithm (see \S\ref{subsec:lenscosts}). For any lens $\Lens$, we can calculate
the expected number of bits required to recover a string $t$ in
$\LanguageOf{\RegexAlt}$ from a synchronized string $s$ in
$\LanguageOf{\Regex}$. This expectation is the \emph{conditional entropy of
  $\Regex$ given $\RegexAlt$ and $\Lens$}, formally $\sum_{s \in
  \Regex}\ProbabilityOf{\Regex}{s}\cdot\EntropyOf{\SetOf{t \SuchThat
    \Lens.\PutROf{s}{t} = t}}$. The likelihood we assign to $\Lens$ is the sum
of the conditional entropy of $\Regex$ given $\RegexAlt$ and $\Lens$ and the
conditional entropy of $\RegexAlt$ given $\Regex$ and $\Lens$. This metric
assigns higher entropy (or cost\bcp{= lower likelihood?  I'm still confused
  by these different ways of speaking about what seems to be the same
  thing.}) to lenses where knowing the string on one side 
provides little information about the string on the other side. It assigns zero
entropy to bijections because given a string $s \in \Regex$, the bijection
exactly determines the corresponding string in $\RegexAlt$.

@@

To obtain SREs from the plain regular expressions that users write, 
we use a default heuristic that attempts to assign probability annotations
giving
each string in the language equal probablity. In other words, this heuristic
tries to make each bit of data no more or less likely than any
other.\bcp{huh?} 
Alternatively, our synthesis algorithm could take as input a pair of stochastic
regular expressions rather than plain ones, permitting users to specify a
data-specific probability distribution (either manually calculated or inferred
from some dataset of interest).\bcp{So why didn't we do it this way?}

However\bcp{``However'' means that something from before is being
  contradicted.  What is it in this case?}, sometimes users already know that certain data should or should not be
used to construct the other format. We introduce relevance annotations to SREs
that enable users to specify whether a piece of data should be used to construct
the other format (with \SRequire) or not (with \Skip). In our example,
\lstinline{salary} and \lstinline{company} could safely be skipped (as they are
disconnected), where \lstinline{name} in \lstinline{emp_salary} and
\lstinline{emp_ins} be annotated as required (as they are converted with the
identity lens).  \bcp{That paragraph seems way too brief and high level.
  These annotations were added in response to significant complaints from an
earlier PC, but the paper reads as though they are an afterthought of little
consequence; they deserve more attention and explanation.}

@@

This architecture of communicating synthesizers gives us a way to enumerate
pairs of stochastic regular expressions of increasing rank and to efficiently
search through them, but it poses a problem: when should \Expand stop proposing
new SRE pairs? We might have found a promising lens between a pair of stochastic
regular expressions, but a different pair we haven't yet discovered may give
rise to an even better lens. The search algorithm must resolve a tension between
the quality of the inferred lens and the amount of time it takes to return a
result. For example, if the algorithm has already found the lens in
Figure~\ref{fig:example_lens}, we don't want to spend a lot of time searching
for an even better lens. To resolve this tension, the algorithm uses heuristics
to judge whether to return the current best satisfying lens to the user or to
pass the next set of unrollings to \GreedySynth. The heuristics favor stopping
if the current best satisfying lens is very likely, indicating the lens is very
promising (for example, if the satisfying lens loses no information, the
algorithm should terminate for no other lens will be more likely\bcp{this
  case is 100\% obvious, which makes it not so interesting as an example.
  The reader is left wondering what happens with pretty likely but not
  maximally likely lenses?}). The
heuristics also favor stopping if \Expand has delivered to \GreedySynth all
pairs of stochastic regular expressions at a given rank and there is a large
number of pairs at the next rank, because searching through all such pairs will
take a long time. With this approach, the algorithm can quickly respond with a
satisfying lens with relatively high likelihood. If the user is unhappy with the
result, they can either refine the search by supplying additional examples,
which serve both to rule out previously proposed lenses and to reduce the
size of the search space by cutting down on the number of satisfying lenses, or
they can supply annotations on the source and target SREs indicating that
certain information either must be retained or must be discarded by the lens
(see \S\ref{subsec:optimizations}). \bcp{Again, I feel like we're
  downplaying this aspect too much.}

@@

\subsection{Lens Costs}
\label{subsec:lenscosts} \bcp{In the intro, we talked about ``likelihood.''
Now we seem to be switching the terminology to ``cost.''}
Our cost metric is based on the expected amount of information required to
recover a string in one data format from the other. We use the function
$\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ to calculate bounds on the
expected amount of information required to recover a string in $\RegexAlt$ from
a string in $\Regex$, synchronized by $\Lens$. Similarly, we use the function
$\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ to calculate bounds on the
expected amount of information required to recover a string in $\Regex$ from a
string in $\RegexAlt$, synchronized by $\Lens$. We write $a(b,c)$ to
mean $(a*b,a*c)$, and $(a,b)+(c,d)$ to mean $(a+c,b+d)$.
\begin{center}
  \begin{tabular}{rcl}
    $\REntropyOf{\RegexAlt \Given \IdentityLensOf{\RegexAlt}, \RegexAlt}$
    & =
    & [0,0]\\

@@

The cost of a lens between two SREs $cost(\Lens, \Regex, \RegexAlt) =
\MaxOf{\LEntropyOf{\Regex \Given \Lens, \RegexAlt}} +
\MaxOf{\REntropyOf{\RegexAlt \Given \Lens, \Regex}}$ is the sum of the maximum
expected number of bits to
recover the left format from the right, and the right from the left. We have
proven theorems demonstrating the calculated entropy corresponds to the actual
conditional entropy to recover the data.

@@

To summarize, the input regular
expressions are first converted into stochastic regular expressions with
\ToStochastic. This pair of SREs is used to initialize a priority queue ($pq$).
The priority of a SRE pair is the number of rewrites needed to derive the pair
from the originals. Next, $\PCF{SynthSymLens}$ enters a loop that searches for
likely lenses. The loop terminates when the algorithm believes it is unlikely to
find a better lens than the best one it has found so far (a termination
condition defined by $\PCF{\Continue}$). Within each iteration of the loop, it:
\begin{itemize}
\item pops the next class ($S$, $T$) of lenses to
search off of the priority queue ($\PCF{\PQ.Pop}$),
\item executes $\GreedySynth$ to find a best lens in that class if
one exists ($\LC$),
using the examples $exs$ to filter out potential lenses that do not satisfy
the specification,
\item replaces $best$ with $\LC$, if $\LC$ has lower cost according
to our information-theoretic metric, and
\item adds the SREs derived from rewriting $S$ and $T$ ($\Expand(S,T)$)
  to the priority queue.
\end{itemize}
When the loop terminates, the search returns the globally best lens found
($best$).  Each subroutine of this algorithm will be explained in
further depth in the following subsections.

\begin{algorithm}[t]
  \caption{\SynthSymLens}
  \label{alg:synth-sym-lens}
  \begin{algorithmic}[1]
    \Function{SynthSymLens}{$\BRegex,\BRegexAlt,\Examples$}
    \State $\Regex \gets \Call{\ToStochastic}{\BRegex}$
    \State $\RegexAlt \gets \Call{\ToStochastic}{\BRegexAlt}$
    \State $\RXSearchState \gets \Call{\PQ.Create}{\Regex,\RegexAlt}$
    \State $\Best \gets \None$
    \While{$\Call{\Continue}{\RXSearchState,\Best}$}
    \State $(\Regex,\RegexAlt) \gets \Call{\PQ.Pop}{\RXSearchState}$
    \State $\LC \gets \Call{\GreedySynth}{\Examples,\Regex,\RegexAlt}$
    \If{$\Call{Cost}{\LC} < \Call{Cost}{\Best}$}
    \State $\Best \gets \LC$
    \EndIf
    \State $\Call{\PQ.Push}{\RXSearchState,\Expand(\Regex,\RegexAlt)}$
    \EndWhile
    \State \ReturnVal{best}
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

\subsection{Relevance Annotations}

@@

Similarly, users can recognize that employee names are present in both files. By
annotating instances of \lstinline{name} as \lstinline{require(name)}, users can
add this knowledge to the specification to optimize the search and force the
the generated lens to retain \lstinline{name} information. In practice, we
define the cost of any lens that loses ``required'' information to be infinite.
%
\begin{center}
  \begin{tabular}{rcl}
    $\REntropyOf{\SRequireOf{\RegexAlt} \Given \Lens, \Regex}$
    & =
    & $\begin{cases}
      \infty & \text{if }\REntropyOf{\SRequireOf{\RegexAlt} \Given \Lens, \Regex}
      \neq 0\\
    0 & \text{otherwise}
  \end{cases}$\\

@@

\bcp{Future work??????????????}

>>>>

In summary, our contributions are as follows:
\begin{itemize}
%
\item We define {\em simple symmetric lenses}, a subclass of symmetric
lenses with no hidden state,
%
we propose a syntax for them---a collection
of combinators for building simple symmetric lenses---and we
describe how to calculate their likelihoods
(\S\ref{sec:overview}--\S\ref{sec:ssl}).  
\item We give an algorithm for synthesizing lenses built from these combinators,
  using a novel application of stochastic regular expressions to guide the
  search process (\S\ref{sec:synthesis}) and prove that our metrics correspond
  to the information loss of a given lens.
%
\item We extend the Boomerang implementation with simple symmetric lenses
and lens synthesis specifications, 
  including types, examples, and relevance annotations. We evaluate our
  implementation, the effects of
  optimizations, and our inclusion of relevance annotations on a set of
  48 lens synthesis benchmarks drawn from data cleaning, view maintenance, and
  file synchronization tasks. The system can synthesize simple symmetric lenses
  for all of the benchmarks in under 30 seconds (\S\ref{sec:evaluation}).
%
\item We prove that simple symmetric lenses fall between asymmetric
lenses~\cite{Focal2005-long} and full symmetric
lenses~\cite{symmetric-lenses} in expressiveness.  Indeed, the
class of simple lenses can be characterized semantically as the subset of
full symmetric lenses for which a simple ``forgetfulness'' property holds
(\S\ref{sec:related}).
\end{itemize}
Along the
way, we show that {\em star-semiring equivalences}~\cite{starsemiring} on regular
expressions 
remain valid when extended to stochastic regular expressions, preserving
probability distributions as well as languages (\S\ref{sec:sre}).
We close with related work (\S\ref{sec:related}) and
concluding thoughts (\S\ref{sec:conc}).
Elided proofs can be found in the appendices
\ifappendices\else included with the supplemental material\fi.

@@

\subsection{Simple Symmetric Lenses}
Semantically, a simple symmetric lens (we will just say ``lens'' when clear
from context) between 
sets $X$ and $Y$ comprises four functions subject to four round-tripping
laws.
\vspace*{.7ex}
\begin{trivlist}
\item 
%\begin{enumerate}
%\item $\CreateR \OfType \Regex \rightarrow \RegexAlt$
%\item $\CreateL \OfType \RegexAlt \rightarrow \Regex$
%\item $\PutR \OfType \Regex \rightarrow \RegexAlt \rightarrow \RegexAlt$
%\item $\PutL \OfType \RegexAlt \rightarrow \Regex \rightarrow \Regex$
%\end{enumerate}
\begin{center}
\begin{minipage}[c]{.3\textwidth}
\begin{gather}
 \mbox{$\CreateR \OfType X \rightarrow Y$} \nonumber \\
 \mbox{$\CreateL \OfType Y \rightarrow X$} \nonumber \\
 \mbox{$\PutR \OfType X \rightarrow Y \rightarrow Y$} \nonumber \\
 \mbox{$\PutL \OfType Y \rightarrow X \rightarrow X$} \nonumber
\end{gather} 
\end{minipage} 
\hspace{.4in}
\begin{minipage}[c]{.5\textwidth}
\begin{align}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{x})}{x} = x\\
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{y})}{y} = y\\
  \tag{\PutRL}
  \PutLOf{(\PutROf{x}{y})}{x} = x\\
  \tag{\PutLR}
  \PutROf{(\PutLOf{y}{x})}{y} = y
\end{align}
\end{minipage} 
\end{center}
\end{trivlist}
The two \KW{create} functions are used to fill 
in default values when introducing new data (\EG, on \KW{create} the ``unk''
salary entry is added alongside the name to the management file when HR inserts
a new employee). The two \KW{put} functions propagate edits from one format to
the other by combining a new value from one with an old value from the other.
The record projection notation $\Lens.\PutR$ extracts the \PutR function
from the 
lens $\Lens$. These four functions can be used to keep the
common information between two file types in sync. For example, if a new file
of the left-hand format is created, the \CreateR function will build a new
file in the 
right-hand format. If the right-hand file is then edited, the \PutL function can
update the left-hand file with the changed information.

@@

% concat(ins(header),employees_lens) : emp_salaries $\Leftrightarrow$ emp_insurance
%
%\ksf{I don't think this example is necessary.}
%Data can also appear in multiple forms. For example,
%a list of employees with insurance information can either be empty or
%be a sequence of employee/insurance records: \ksf{This example is
%vacuous as you could eliminate the or in favor of the *.  Do somethign
%else to introduce or?}
%
%\begin{lstlisting}
%or(id(""),employees_lens) : emp_salaries $\Leftrightarrow$ "" | ("\n" emp_ins) . ("\n" emp_ins)*
%\end{lstlisting}
The \IterateLens lens is useful for synchronizing a series of items or rows in a
table. For example given a
lens \lstinline{employee_lens} that synchronizes data for a single employee, the
lens
%
\begin{lstlisting}
iterate(id("\n") . employee_lens)  :  ("\n" . emp_salary)$^*$ $\Leftrightarrow$ ("\n" . emp_ins)$^*$
\end{lstlisting}
%
transforms a list of employees in employees in the Management format to a list
of employees in the HR format and vice versa. These combinators
are combined in figure~\ref{fig:example_lens} to construct a complete lens
between the employee formats.

@@

\bcp{I wonder whether the rest of this subsection can be cut... It is both
  hard to understand and too brief to really say much, and I wonder whether
  we need it for anything between here and where we give it a real
  explanation.} \afm{I believe it is necessary for 2.4. I tried giving it a more
terse treatment, but that caused its own set of problems.}

Lenses are typed by pairs of regular expressions and only a handful of
rules result in a given type.  For example, the type of a
\ConcatLens lens is the concatenation of the regular expressions of the sub-lenses.
\[
  \inferrule*
  {
    \Lens_1 \OfType \BRegex_1 \Leftrightarrow \BRegexAlt_1\\
    \Lens_2 \OfType \BRegex_2 \Leftrightarrow \BRegexAlt_2
  }
  {
    \ConcatLensOf{\Lens_1}{\Lens_2} \OfType \BRegex_1 \Concat \BRegex_2
    \Leftrightarrow
    \BRegexAlt_1 \Concat \BRegexAlt_2
  }
\]
The typing  rules for maintain structural similarity
between regular expression types and lens terms, with one exception: The
type equivalence rule permits a well-typed lens to be well-typed among any
equivalent regular expression pair.
%
\[
  \centering
  \inferrule*
  {
    \Lens \OfType \BRegex \Leftrightarrow \BRegexAlt\\
    \BRegex \SSREquiv \BRegex'\\
    \BRegexAlt \SSREquiv \BRegexAlt'
  }
  {
    \Lens \OfType \BRegex' \Leftrightarrow \BRegexAlt'
  }
\]
%
This rule is needed for lenses like the final one in
Figure~\ref{fig:example_lens} to be well-typed.

@@

One challenge is that the simple symmetric lens combinators permit many
well-typed lenses
between a given pair of regular expressions. For example, 
Figure~\ref{fig:example_lens} gives one possible lens between regular
expressions  \lstinline{emp_salaries} and \lstinline{emp_insurance},
but 
%
\begin{lstlisting}
disconnect(emp_salaries, emp_insurance, ..., ...)
\end{lstlisting}
%
is another well-typed lens that satisfies the example in
Figure~\ref{fig:minimized-representations} (where \lstinline{...} represents
the example strings). In general, {\em many} examples
may be 
required to rule out all possible occurrences of \Disconnect lenses,
particularly in complex formats.
%\ksf{I don't understand the following.  We shouldn't explain the issue in
%terms of \PutR and \PutL as we are focused on the combinators rather
%than the raw functions.}
%The default strings given to \Disconnect merely need to be the strings
%in the input-output examples, and the lens will still satisfy the
%specification.  While the disconnect lens is prohibited by two
%examples, or by a carefully chosen input-output example on \PutR
%or \PutL, this problem continues presenting itself for the sublenses,
%and is highlighted even more on larger formats.
Instead of merely finding
\emph{any} satisfying lens, we wish to synthesize a satisfying lens that is
likely to please the user.

How can we identify such a ``likely'' lens? We propose the following heuristic:
A satisfying lens is ``more likely'' if it uses more data from one format to
construct the other.  For example, the
identity lens (which uses all the data) is more likely than the disconnect
lens (which uses none). Formally, we define the {\em likelihood} of a
satisfying lens as the 
expected number of bits required to recover data in one format from 
data in the other; higher likelihoods correspond to fewer bits. Two
strings $s$ and $t$ are \emph{synchronized according to lens
  $\Lens$} if $\Lens.\PutR\
s\ t = t$ and $\Lens.\PutL\ t\ s = s$. We can \emph{recover} $s$ from 
$t$ using bits $b$ and lens $\Lens$ if we can reconstruct $s$ from $t$, $b$, and
$\Lens$. For example, given the \IdentityLens{} lens, we can recover $s$ from $t$
using no bits because, in this case, $s$ is just $t$. In contrast, given the
\Disconnect{} lens, we need enough bits to fully encode $s$ in order to recover
it from $t$ because all the information in $t$ gets thrown away. Thus, if both
\IdentityLens{} and \Disconnect{} are satisfying lenses for a particular pair of
formats, \IdentityLens{} will be more likely.

@@

We use entropy to gauge the relative likelihood of lenses in our synthesis
algorithm (see \S\ref{subsec:lenscosts}). For any lens $\Lens$, we can calculate
the expected number of bits required to recover a string $t$ in
$\LanguageOf{\RegexAlt}$ from a synchronized string $s$ in
$\LanguageOf{\Regex}$. This expectation is the \emph{conditional entropy of
  $\Regex$ given $\RegexAlt$ and $\Lens$}, formally $\sum_{s \in
  \Regex}\ProbabilityOf{\Regex}{s}\cdot\EntropyOf{\SetOf{t \SuchThat
    \Lens.\PutROf{s}{t} = t}}$. The likelihood we assign to $\Lens$ is the sum
of the conditional entropy of $\Regex$ given $\RegexAlt$ and $\Lens$ and the
conditional entropy of $\RegexAlt$ given $\Regex$ and $\Lens$. This metric
assigns higher entropy (or likelihood) to lenses where knowing the string on one side 
provides little information about the string on the other side. It assigns zero
entropy to bijections because given a string $s \in \Regex$, the bijection
exactly determines the corresponding string in $\RegexAlt$.

@@

To obtain SREs from the plain regular expressions that users write, 
we use a default heuristic that attempts to assign probability annotations
giving
each string in the language equal probablity (not prioritizing one piece of
information over another).

Sometimes users already know that certain data should or should not be
used to construct the other format. We introduce relevance annotations to SREs
that enable users to specify whether a piece of data should be used to construct
the other format (with \SRequire) or not (with \Skip). In our example,
\lstinline{salary} and \lstinline{company} could safely be skipped (as they are
disconnected), where \lstinline{name} in \lstinline{emp_salary} and
\lstinline{emp_ins} be annotated as required (as they are converted with the
identity lens). Doing so both constrains the problem (as \lstinline{name}s must
be synchronized) and makes it easier (as the algorithm does not waste time
looking for \lstinline{salary} information in the insurance format). In this
way, users can tweak the
lens likelihoods with their external knowledge.

@@

This architecture of communicating synthesizers gives us a way to enumerate
pairs of stochastic regular expressions of increasing rank and to efficiently
search through them, but it poses a problem: when should \Expand stop proposing
new SRE pairs? We might have found a promising lens between a pair of stochastic
regular expressions, but a different pair we haven't yet discovered may give
rise to an even better lens. The search algorithm must resolve a tension between
the quality of the inferred lens and the amount of time it takes to return a
result. For example, if the algorithm has already found the lens in
Figure~\ref{fig:example_lens}, we don't want to spend a lot of time searching
for an even better lens. To resolve this tension, the algorithm uses heuristics
to judge whether to return the current best satisfying lens to the user or to
pass the next set of unrollings to \GreedySynth. The heuristics favor stopping
if the current best satisfying lens is very likely, indicating the lens is very
promising (for example, if the satisfying lens loses no information, the
algorithm should terminate for no other lens will be more likely). The
heuristics also favor stopping if \Expand has delivered to \GreedySynth all
pairs of stochastic regular expressions at a given rank and there is a large
number of pairs at the next rank, because searching through all such pairs will
take a long time (for example, if the satisfying lens loses only a little
information,
but searching for a better one will take a long time, the discovered lens is
returned).
With this approach, the algorithm can quickly return a
satisfying lens with relatively high likelihood. If the user is unhappy with the
result, they can either refine the search by supplying additional examples,
which serve both to rule out previously proposed lenses and to reduce the
size of the search space by cutting down on the number of satisfying lenses, or
they can supply annotations on the source and target SREs indicating that
certain information either must be retained or must be discarded by the lens
(see \S\ref{subsec:relevanceannotations}).

@@

\subsection{Lens Likelihoods}
\label{subsec:lenscosts}
Our likelihood metric is based on the expected amount of information required to
recover a string in one data format from the other. We use the function
$\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ to calculate bounds on the
expected amount of information required to recover a string in $\RegexAlt$ from
a string in $\Regex$, synchronized by $\Lens$. Similarly, we use the function
$\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ to calculate bounds on the
expected amount of information required to recover a string in $\Regex$ from a
string in $\RegexAlt$, synchronized by $\Lens$. We write $a(b,c)$ to
mean $(a*b,a*c)$, and $(a,b)+(c,d)$ to mean $(a+c,b+d)$.
\begin{center}
  \begin{tabular}{rcl}
    $\REntropyOf{\RegexAlt \Given \IdentityLensOf{\RegexAlt}, \RegexAlt}$
    & =
    & [0,0]\\

@@

The likelihood of a lens is the negative of its \emph{cost}. The cost of a lens
between two SREs $cost(\Lens, \Regex, \RegexAlt) = \MaxOf{\LEntropyOf{\Regex
    \Given \Lens, \RegexAlt}} + \MaxOf{\REntropyOf{\RegexAlt \Given \Lens,
    \Regex}}$ is the sum of the maximum expected number of bits to recover the
left format from the right, and the right from the left. We have proven theorems
demonstrating the calculated entropy corresponds to the actual conditional
entropy to recover the data.

@@

To summarize, the input regular
expressions are first converted into stochastic regular expressions with
\ToStochastic. This pair of SREs is used to initialize a priority queue ($pq$).
The priority of a SRE pair is the number of rewrites needed to derive the pair
from the originals. Next, $\PCF{SynthSymLens}$ enters a loop that searches for
likely lenses. The loop terminates when the algorithm believes it is unlikely to
find a better lens than the best one it has found so far (a termination
condition defined by $\PCF{\Continue}$). Within each iteration of the loop, it:
\begin{itemize}
\item pops the next class ($S$, $T$) of lenses to
search off of the priority queue ($\PCF{\PQ.Pop}$),
\item executes $\GreedySynth$ to find a best lens in that class if
one exists ($\Lens$),
using the examples $exs$ to filter out potential lenses that do not satisfy
the specification,
\item replaces $best$ with $\Lens$, if $\Lens$ is more likely according
to our information-theoretic metric, and
\item adds the SREs derived from rewriting $S$ and $T$ ($\Expand(S,T)$)
  to the priority queue.
\end{itemize}
When the loop terminates, the search returns the globally best lens found
($best$).  Each subroutine of this algorithm will be explained in
further depth in the following subsections.

\begin{algorithm}[t]
  \caption{\SynthSymLens}
  \label{alg:synth-sym-lens}
  \begin{algorithmic}[1]
    \Function{SynthSymLens}{$\BRegex,\BRegexAlt,\Examples$}
    \State $\Regex \gets \Call{\ToStochastic}{\BRegex}$
    \State $\RegexAlt \gets \Call{\ToStochastic}{\BRegexAlt}$
    \State $\RXSearchState \gets \Call{\PQ.Create}{\Regex,\RegexAlt}$
    \State $\Best \gets \None$
    \While{$\Call{\Continue}{\RXSearchState,\Best}$}
    \State $(\Regex,\RegexAlt) \gets \Call{\PQ.Pop}{\RXSearchState}$
    \State $\Lens \gets \Call{\GreedySynth}{\Examples,\Regex,\RegexAlt}$
    \If{$\Call{Cost}{\Lens} < \Call{Cost}{\Best}$}
    \State $\Best \gets \Lens$
    \EndIf
    \State $\Call{\PQ.Push}{\RXSearchState,\Expand(\Regex,\RegexAlt)}$
    \EndWhile
    \State \ReturnVal{best}
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

\subsection{Relevance Annotations}
\label{subsec:relevanceannotations}

@@

Similarly, users can recognize that employee names are present in both files. By
annotating instances of \lstinline{name} as \lstinline{require(name)}, users can
add this knowledge to the specification to optimize the search and force the
the generated lens to retain \lstinline{name} information. In practice, we
make any lens that loses ``required'' information infinitely unlikely.
%
\begin{center}
  \begin{tabular}{rcl}
    $\REntropyOf{\SRequireOf{\RegexAlt} \Given \Lens, \Regex}$
    & =
    & $\begin{cases}
      \infty & \text{if }\REntropyOf{\SRequireOf{\RegexAlt} \Given \Lens, \Regex}
      \neq 0\\
    0 & \text{otherwise}
  \end{cases}$\\

@@

\bcp{Future work??????????????}\afm{I don't think there's room for it}
